### 2011-12-22 tr-semanticsearchsurvey-data/mr.tex

\section{Matching \& Ranking in Semantic Data Retrieval}

\subsection{Query-Independent Ranking}
- Existing approaches based on PageRank

\subsection{Term-based Matching \& Ranking}
- Does not consider structure information \\
- Most existing engines use simple TF/IDF term weighting, do not take structure information into account

\subsection{Term-based Matching \& Structured-based Ranking}
- Take structure of data \& results into account for ranking\\
-	BM25F\\
- Language model for structured results

\subsection{Term- and Structure-based Matching \& Ranking} 
-	Matching: take structure of data account for computing structured results