### 2011-12-22 tr-semanticsearchsurvey-data/sdt.tex

\section{Semantic Data Retrieval}
\subsection{Applications} 
\subsection{Semantic Data Retrieval Process} 
- Crawling, integration, indexing, matching \& ranking, result presentation \\

\subsection{Taxonomy of Semantic Data Retrieval Approaches}
- overview of existing approaches categorized along the dimensions: need, data, queries, targeted task in the process \\
- following survey focus on keyword queries  \\

\subsection{Semantic Data Crawling} 
\subsection{Semantic Data Integration}
\subsection{Semantic Data Storage and Indexing}
\subsection{Semantic Search Result Presentation}