### 2012-05-10 paper-iswc-reification/paper.tex

\documentclass{llncs}
%
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{epsfig}
\usepackage{url}
\usepackage{color}
\usepackage[utf8]{inputenc}
%\usepackage{subfig}
%\usepackage{algorithm}
%\usepackage{algorithmic}
\usepackage{makeidx} % allows for indexgeneration
%\usepackage{natbib}
\usepackage{amsmath,amsfonts,amssymb}
%\usepackage[top=3.95cm, bottom=4cm, left=4.6cm, right=4.6cm]{geometry}

% \usepackage{latexsym}
% \usepackage{amsmath}
% \usepackage{amssymb}
\newcommand{\dtr}[1]{(\textbf{Thanh: #1})}
%\newcommand{\todo}[1]{
%\textbf{(todo: #1)}
%}

%\input{commands}

%\renewcommand{\baselinestretch}{0.98}
\frontmatter          % for the preliminaries
\pagestyle{headings}  % switches on printing of running heads
\mainmatter           % start of the contributions

%\sloppy
\title{Efficient Management of Meta-level Information for RDF}

%\thanks{This work has been supported by the European Commission
%under contract IST-2006-027595 NeOn, and by the Deutsche
%Forschungsgemeinschaft (DFG) under the ReaSem project.}}

%\titlerunning{}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Olaf Hartig$^1$, Mike Personick$^2$, Yrjana Rankka$^3$, Bryan Thompson$^2$, Thanh Tran$^4$}

\authorrunning{Hartig}   % abbreviated author list (for running head)

%%% modified list of authors for the TOC (add the affiliations)
\tocauthor{}

\institute{
$^1$Humboldt University, Berlin, Germany\\
$^2$SYSTAP LLC, Washington, USA\\
$^3$OpenLink Software, Burlington, USA\\
$^4$Institute AIFB, Karlsruhe Institute of Technology, Germany
}

\makeindex

\begin{document}

\maketitle

\begin{abstract}
\end{abstract}

\input{intro}
\input{related}
\input{rep}
\input{querying}
\input{dis}
\input{exp}
\input{conc}

%\begin{appendix}
%\input{appendix}
%\end{appendix}
%
\bibliographystyle{abbrv}
\bibliography{paper}
\end{document}