### 2012-05-10 paper-iswc-reification/rep.tex

\section{Representation}
Syntax 
Expressed tersely as:
  <<:SAP :bought :sybase>> dc:source reuters:us-sybase .
Statements are “wrapped” by << … >> 
Turtle Extension
<<:SAP :bought :sybase>> dc:source reuters:us-sybase ;
			  dc:created “1995-01-07” ;
                        		 dc:author  “James R. Jones“ .

-	The statement can appear in either the subject or object position.
-	Recursion in statements is allowed
-	But depth is shallow in practice
Semantics