### 2012-05-10 trunk/paper-iswc-reification/dis.tex

\section{Discussion}
What are benefits compared to RDF reification and other related work:
Compact Representation 
Efficient Querying 
 # Who bought Sybase and where did we get that fact?

SELECT ?src ?who {
   <<?who :bought :sybase>> dc:source ?src
}

Answered efficiently
-	2-bound POS index scan (?who :bought sybase) => ?sid
-	Join with (?sid dc:source ?src)

# Explicit bind of statement on variable for multiple joins.

SELECT ?src ?who ?created {
   <<?who :bought :sybase>> as ?sid
   ?sid dc:source ?src
   OPTIONAL {?sid dc:created ?created}
}

Compatibility with Existing Standards
-	Harmonized with RDF Reification
-	Can be combined with named graphs