### 2012-05-10 trunk/paper-iswc-reification/exp.tex

\section{Experiments}
Show the benefits in experiments
Systems
-	Bigdata
-	Virtuoso

Data
-	Datasets + artificially generated meta-level statements
-	Size of the data
-	Size of the indexes 
Loading Time Performance
Querying Time Performance