### 2012-05-10 trunk/paper-iswc-reification/intro.tex

\section{Introduction}
Motivation
\begin{itemize}
	\item 
Necessary in several domains
\item	Entity link extraction (confidence values)
\item	Graph databases (link attributes)
\item	Intelligence analysis (provenance on assertions)
\item	Statement level security
\end{itemize}