### 2012-05-13 trunk/paper-iswc-reification/commands.tex
added 'commands.tex'

%%% review commands:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcounter{todo}

\newcommand{\todo}[1]
	{\stepcounter{todo}\noindent{\color{red} \textsf{Todo~{\arabic{todo}}: #1}}}  % use this for drafts
	%{}                                                                           % use this for the final version

\newcommand{\Todo}[2]
	{{\color{red}\textsf{#1}\footnote{\todo{#2}}}}  % use this for drafts
	%{#1}                                           % use this for the final version

\newcommand{\removable}[1]%
	{{\color{gray}#1}}  % use this for drafts
	%{#1}               % use this for the final version

%%% key terminology:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\RDFplus}{RDF$^{\mathsf{\text{\tiny +\!meta}}}$}
\newcommand{\RDFplusTriple}{{\RDFplus} triple}

\newcommand{\TRef}%
	%{triple reference}
	{tref}
\newcommand{\TRefLongVersion}{triple reference}

%%% relevant math symbols
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\renewcommand{\emptyset}{\varnothing}

\newcommand{\symAllURIs}{\mathcal{U}} % the symbol for the set of all URIs
\newcommand{\symAllLiterals}{\mathcal{L}} % the symbol for the set of all literals
\newcommand{\symAllBNodes}{\mathcal{B}} % the symbol for the set of all blank nodes
\newcommand{\symAllVariables}{\mathcal{V}} % the symbol for the set of all query variables
	\newcommand{\symAllTRefsInfinite}{\mathcal{R}} % don't use that; it was required to define \symAllTRefs
\newcommand{\symAllTRefs}{\overline{\symAllTRefsInfinite}} % the symbol for the set of all triple references

\newcommand{\symTermsFct}{\mathrm{terms}} % the name used for \fctTerms
\newcommand{\fctTerms}[1]{\symTermsFct(#1)} % the set of all terms in triple/graph #1

\newcommand{\symTRefFct}{\mathrm{trefs}} % the name used for \fctTref
\newcommand{\fctTRef}[1]{\symTRefFct(#1)} % the set of all triple references in triple/graph #1

\newcommand{\symDerefFct}{\mathit{deref}} % the name used for \fctDeref
\newcommand{\fctDeref}[1]{\symDerefFct(#1)} % the triple denoted by triple reference #1

\newcommand{\mathTinyRDF}[1]{\texttt{\footnotesize #1}}