### 2012-05-28 trunk/paper-iswc-reification/dis.tex

\section{Discussion}
What are benefits compared to RDF reification and other related work. We discuss the benefits in terms of data management and query processing.

\subsection{Data Management}
- model efficient representation: less triples needed\\
- thus more compact serialization: smaller turtle export\\

\subsection{Query Processing}
- more efficient processing because less data has to be loaded and \\
- less joins are needed to combine binding for different triple patterns\\

 \# Who bought Sybase and where did we get that fact?

SELECT ?src ?who {
   <<?who :bought :sybase>> dc:source ?src
}

Answered efficiently
-	2-bound POS index scan (?who :bought sybase) => ?sid
-	Join with (?sid dc:source ?src)

\# Explicit bind of statement on variable for multiple joins.

SELECT ?src ?who ?created {
   <<?who :bought :sybase>> as ?sid
   ?sid dc:source ?src
   OPTIONAL {?sid dc:created ?created}
}

Compatibility with Existing Standards
-	Harmonized with RDF Reification
-	Can be combined with named graphs