### 2012-06-02 paper-iswc-reification/paper.tex

\documentclass{llncs}

\usepackage{amsmath,amssymb}
% \usepackage[ruled,vlined]{algorithm2e}
\usepackage{color}  \definecolor{gray}{rgb}{.5,.5,.5}
\usepackage{times}
% \usepackage{graphicx}
% \usepackage{subfig}
% \usepackage{epsfig}
% \usepackage{url}
% \usepackage[utf8]{inputenc}
% \usepackage{paralist}
%\usepackage{subfig}
%\usepackage{algorithm}
%\usepackage{algorithmic}
% \usepackage{makeidx} % allows for indexgeneration
%\usepackage{natbib}
%\usepackage[top=3.95cm, bottom=4cm, left=4.6cm, right=4.6cm]{geometry}

% \usepackage{latexsym}

\input{commands}

%\renewcommand{\baselinestretch}{0.98}
% \frontmatter          % for the preliminaries
% \pagestyle{headings}  % switches on printing of running heads
% \mainmatter           % start of the contributions

%\sloppy
\title{Efficient Management of Reified Statements in RDF}

%\thanks{This work has been supported by the European Commission
%under contract IST-2006-027595 NeOn, and by the Deutsche
%Forschungsgemeinschaft (DFG) under the ReaSem project.}}

%\titlerunning{}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Olaf Hartig\inst{1} \and Mike Personick\inst{2} \and Yrj{\"a}n{\"a} Rankka\inst{3} \and Bryan Thompson\inst{2} \and Thanh Tran\inst{4}}

\authorrunning{Hartig}   % abbreviated author list (for running head)

%%% modified list of authors for the TOC (add the affiliations)
\tocauthor{}

\institute{%
Humboldt-Universit{\"a}t zu Berlin, Berlin, Germany\\
\and
SYSTAP LLC, Washington, USA\\
\and
OpenLink Software, Burlington, USA\\
\and
Institute AIFB, Karlsruhe Institute of Technology, Germany
}

\makeindex

\begin{document}

\maketitle

\begin{abstract}
Metadata is commonly associated with RDF statements to capture provenance as well as quality related aspects. RDF reification is the standard mechanism to express this kind of information at the statement-level. In this work, we revisit the drawbacks of this mechanism. Specifically in the RDF data management and query processing context, we show that not only it results in a large number of triples but also, requires a large number of joins in queries asking for RDF statements' metadata. Addressing these problems, we propose a solution for reification that extends the current RDF and SPARQL standards. Based on two commercial RDF stores adoption of this solution, we show in experiments that compared to the original RDF reification proposal, it significantly improves the efficiency of RDF data management and SPARQL query processing. 
\end{abstract}

\input{intro}
\input{related}
\input{dis}
\input{rep}
\input{querying}
\input{exp}
\input{conc}

%\begin{appendix}
%\input{appendix}
%\end{appendix}
%
\bibliographystyle{abbrv}
\bibliography{paper}
\end{document}