### 2012-06-07 trunk/paper-iswc-reification/mainext.tex
added a second version of the paper that is meant to become the extended version with an appendix that contains all proofs

\documentclass{llncs}

\usepackage{amsmath,amssymb}
% \usepackage[ruled,vlined]{algorithm2e}
\usepackage{color}  \definecolor{gray}{rgb}{.5,.5,.5}
\usepackage{times}
% \usepackage{graphicx}
% \usepackage{subfig}
% \usepackage{epsfig}
% \usepackage{url}
% \usepackage[utf8]{inputenc}
% \usepackage{paralist}
%\usepackage{subfig}
%\usepackage{algorithm}
%\usepackage{algorithmic}
% \usepackage{makeidx} % allows for indexgeneration
%\usepackage{natbib}
%\usepackage[top=3.95cm, bottom=4cm, left=4.6cm, right=4.6cm]{geometry}

% \usepackage{latexsym}

\input{commands}

%\renewcommand{\baselinestretch}{0.98}
% \frontmatter          % for the preliminaries
% \pagestyle{headings}  % switches on printing of running heads
% \mainmatter           % start of the contributions

%\sloppy
\title{Efficient Management of Reified Statements in RDF}
\subtitle{(Extended Version)\footnote{This report presents an extended version of a paper submitted to ISWC 2012. This extended version contains proofs for all
	%propositions, lemmas, and theorems
	technical results
in the paper (cf.~\appendixname~\ref{Appendix:Proofs}).}
% \vspace{-5mm} % Layout Adjustment
}

%\thanks{This work has been supported by the European Commission
%under contract IST-2006-027595 NeOn, and by the Deutsche
%Forschungsgemeinschaft (DFG) under the ReaSem project.}}

%\titlerunning{}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Olaf Hartig\inst{1} \and Mike Personick\inst{2} \and Yrj{\"a}n{\"a} Rankka\inst{3} \and Bryan Thompson\inst{2} \and Thanh Tran\inst{4}}

\authorrunning{Hartig}   % abbreviated author list (for running head)

%%% modified list of authors for the TOC (add the affiliations)
\tocauthor{}

\institute{%
Humboldt-Universit{\"a}t zu Berlin, Berlin, Germany\\
\and
SYSTAP LLC, Washington, USA\\
\and
OpenLink Software, Burlington, USA\\
\and
Institute AIFB, Karlsruhe Institute of Technology, Germany
}

\makeindex

\begin{document}

\maketitle

\begin{abstract}
\input{abstract}
\end{abstract}

\input{intro}
For full technical proofs of all propositions in this paper we refer to \appendixname~\ref{Appendix:Proofs}.

\input{related}
\input{dis}
\input{rep}
\input{querying}
\input{exp}
\input{conc}

%\begin{appendix}
%\input{appendix}
%\end{appendix}
%
\bibliographystyle{abbrv}
\bibliography{paper}

\newpage
\appendix
\input{proofs}

\end{document}