### 2012-06-23 trunk/tr-rankingsurvey/intro.tex

\section{Introduction}\label{sec:intro}

\paragraph{Outline}