### 2012-06-23 tr-rankingsurvey/overview.tex

\section{The Ranking Problem}\label{sec:ranking}

\subsection{Ranking on Structured Data}
Structured data: can be conceived as graphs\\
Ranking: traditionally, core IR problem, many techniques have been adopted to deal with structured data\\
Ranking over structured data: ranked list of results\\
Types of ranked results: entities, relations, subgraphs, entire datasets (graphs)\\
Evaluating ranked results: different metrics\\

\subsection{Ranking on Structured Data vs. Ranking on Unstructured Data}
Discuss differences\\
Say we focus on structured data\\

\subsection{Ranking vs. Matching}
Discuss differences\\
Say we focus on ranking\\

\subsection{Applications} 
Optional / can be short but should provide good coverage of different application domains