### 2012-06-23 tr-rankingsurvey/sample.tex

\section{Sample Approaches from the Literature}
Before, we discuss the concepts/techniques behind existing approaches. Here we provide a sample