### 2012-07-20 tr-rankingsurvey/sample.tex
Abstract, Intro, Ranking problem, Applications and initial list of sample approaches added

\section{Sample Approaches from the Literature}
Before, we discuss the concepts/techniques behind existing approaches. Here we provide a sample

\begin{table*}[htdp]
\begin{center}
\begin{tabular}{|l|l|}
\hline
\textbf{Algorithm}&\textbf{Reference}\\ \hline
OntologyRank & Ding et al., 2004 \cite{Ding2004}\\ \hline
ObjectRank & Balmin, Hristidis, \& Papakonstantinou, 2004 \cite{Balmin2004}\\ \hline
PopRank & Nie, Zhang, Wen \& Ma, 2005 \cite{Nie2005}\\ \hline
SemRank & Anyanwu, Maduko \& Sheth, 2005 \cite{Anyanwu2005}\\ \hline
ReConRank	& Hogan, Harth \& Decker, 2006 \cite{Hogan2006}\\ \hline
AKTiveRank	 & Alani, Brewster \& Shadbolt, 2006 \cite{Alani2006}\\ \hline
Naming Authority Rank & Harth, Kinsella \& Decker, 2009 \cite{Harth2009}\\ \hline
TripleRank & Franz, Schultz, Sizov \& Staab, 2009 \cite{Franz2009}\\ \hline
RareRank & Wei, 2009 \cite{Wei2009}\\ \hline
DBpediaRanker	& Mirizzi, Ragone, Di Noia \& Di Sciascio, 2010 \cite{Mirizzi2010}\\ \hline
DING & Delbru, Toupikov, Catasta, Tummarello \& Decker, 2010 \cite{Delbru2010}\\ \hline
\end{tabular}
\end{center}
\caption{Summary of semantic ranking algorithms and references}
\label{tab:table1}
\end{table*}%

TODO: adapt table to new taxonomy

\begin{table*}[htdp]
\begin{center}
\begin{tabular}{l l l |c|c|c|c|c|c|c|c|c|c|c|} \cline{4-14}
& & & \rotatebox{90}{OntologyRank} & \rotatebox{90}{ObjectRank} & \rotatebox{90}{PopRank} & \rotatebox{90}{SemRank} & \rotatebox{90}{ReConRank} & \rotatebox{90}{AKTiveRank} & \rotatebox{90}{Hart et. al} & \rotatebox{90}{TripleRank} & \rotatebox{90}{RareRank} & \rotatebox{90}{DBpediaRanker} & \rotatebox{90}{DING} \\ \cline{2-14}
& \multicolumn{1}{|c|}{\multirow{2}{*}{Query dependency}} & Static ranking  & x & x & x & x &  &  & x & x & x & x & x\\ \cline{3-14}
								     & \multicolumn{1}{|c|}{} & Dynamic ranking &  & x &  & x & x & x &  &  &  &  & \\ \cline{1-14}
 \multicolumn{1}{|c|}{\multirow{7}{*}{\rotatebox{90}{Granularity}}} & \multicolumn{1}{|c|}{\multirow{5}{*}{Item granularity}} & Document & x &  &  &  &  & x &  &  &  &  & \\ \cline{3-14}
\multicolumn{1}{|c|}{}
														  & \multicolumn{1}{|c|}{} & Dataset &  &  &  &  &  &  & x &  &  &  & x\\ \cline{3-14}
\multicolumn{1}{|c|}{}														  
														  & \multicolumn{1}{|c|}{} & Entity(Resource) &  & x & x &  & x &  &  & x & x & x & x\\ \cline{3-14}
\multicolumn{1}{|c|}{}														  
														  &\multicolumn{1}{|c|}{} & Identifier &  &  &  &  &  &  & x &  &  &  & \\ \cline{3-14}
\multicolumn{1}{|c|}{}														  
														  & \multicolumn{1}{|c|}{} & Relationship &  &  &  & x &  &  &  & x &  &  & \\ \cline{2-14}
						\multicolumn{1}{|c|}{}  &	\multicolumn{1}{|c|}{\multirow{2}{*}{Ranking granularity}} & Document relationships & x &  &  &  &  &  &  &  &  &  & \\ \cline{3-14}						  
						  									\multicolumn{1}{|c|}{}   & \multicolumn{1}{|c|}{} & Entity (Resource) relationships &  & x & x & x & x &  &  & x & x &  & x\\ \cline{1-14}	
& \multicolumn{1}{|c|}{\multirow{5}{*}{Ranking factor}} & Provenance &  & x &  &  & x &  & x & x & x &  & x\\ \cline{3-14}
							     & \multicolumn{1}{|c|}{} & Context (Topic) &  & x & x & x &  & x &  & x & x & x & \\ \cline{3-14}
							     & \multicolumn{1}{|c|}{} & Similarity measure &  &  &  & x &  & x &  &  &  & x & \\ \cline{3-14}
							     & \multicolumn{1}{|c|}{} & Locality & &  &  &  &  &  &  &  &  & & x\\ \cline{3-14}
							     & \multicolumn{1}{|c|}{} & Predictability &  &  &  & x &  &  &  &  &  &  & \\ \cline{2-14}	 																   						  						
\end{tabular}
\end{center}
\caption{Summary of ranking algorithms and data features}
\label{tab:table3}
\end{table*}