### 2012-07-20 trunk/tr-rankingsurvey/future.tex
initial thoughts about open challenges

\section{Open Research Challenges}\label{sec:challenges}

While this survey evidences a considerable degree of progress in ranking for the Web of Linked Data, it also evidences some areas that are still to be developed and several problems that need further research and development. They include the following:
\begin{itemize}
	\item How to add support for data streams. The development of high scalable models is a must to deal with this kind of environments.
	\item Lack of mechanisms to capture the social dimension of linked open data and its exploitation in ranking.
	\item How to include support for personalized ranking functions. If users are the consumers of the information they should be able to customize the consumption.	
\end{itemize}